MSc Thesis

### 2023-06-13 manuscript/main.tex
better graphs and final data

\section{Results}
Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}). We found strong support for a species effect on silicon content ($p=0.012, F = 6.08, df=3,10.3$). Oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}). Counter to our predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species. Despite a small p-value for an induction treatment effect ($p = 0.011, F = 4.62, df =2, 186$), the only significant pairwise differences between treatments were in wheat, where silicon content was lowest in the insect treatment group (Figure \ref{Fig:induction}). We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,186$), likely driven by the stark differences between wheat and the other species' responses to the induction treatments (Figure \ref{Fig:induction}).
We found a significant effect of species, but not induction treatment, on phenolic content (Figure \ref{Fig:induction}, Table \ref{Tab:phe_params}). Oats had the highest average leaf phenolic content [$val \pm err$], while barley had the lowest average phenolic content [$val \pm err$]. 
We did not find a significant relationship between silicon and phenolic content ($p = 0.15$), despite an apparent negative trend in the scatter plot (Figure \ref{Fig:phe_si_scatter}).

@@

We did not see a clear relationship between silicon and phenolic content in our plants, matching the observations of \textcite{waterman_short-term_2021}). Though they observed a negative relationship between silicon and phenolics in plants grown without supplemental silicon, this relationship disappeared when they added silicon to the growth media. As previously discussed, our study seems to have imparted moderated silicon limitation on the plants, and this may explain the hint of a negative trend observable in the scatterplot (Figure \ref{Fig:phe_si_scatter}). Our observations on silicon content variation among species complement those of \textcite{simpson_still_2017}. In their study, they tested 8 species groups of wild and domesticated cereal crops, including wheat and barley. Our results for wheat and barley silicon and phenolic content more or less follow the patterns they observed, with wheat having significantly higher silicon content, and barley having much less, while the two species have more similar phenolic content. \textcite{simpson_still_2017} did not test triticale nor oats, so this study does add first records of the silicon content and phenolic content for these species. Both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we can interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic content among studies and growing environments. In our study, oats had the highest phenolic concentration, and the lowest silicon content. The relative difference in the phenolic and silicon content of oats and wheat in our study [was similar] to the largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study. Though phylogenetic relationships are one driver of silicon and phenolic contents, there are likely proximal evolutionary (eg. domestication) and ecological (eg. growing environment) drivers that also determine the final defensive phenotype of cereal species. Understanding how these various factors constrain the range of defensive expression for these species will help bring focus to efforts to translate silicon ecology into the field of agriculture.

@@

\begin{figure}[h]
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/spp_si_content.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.45\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/spp_phenolic_content.png}
        \end{subfigure}

        \centering
        \caption{Baseline (uninduced) silicon content (\% dry matter)in the cereal cultivars used in this study. Cultivar species is indicated by point color. Error bars are 1 Standard Error.}
        \label{Fig:baseline_si}
\end{figure}

\begin{figure}[h]
        \centering
        \begin{subfigure}[b]{0.45\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/induction_plot_letters.png}
        \end{subfigure}
        \hfill
        \begin{subfigure}[b]{0.45\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/phenolic_induction_plot_letters.png}
        \end{subfigure}
        \caption{The response of leaf silicon content (left) and leaf phenolic content (right) for four cereal species across three defence induction treatments. We treated plants either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}). Leaves were sampled 18 hours after treatment, and were analysed for silicon using XRF or for phenolic content using the Fast Blue BB method. Dots and error bars are the mean $\pm$ 1 Standard Error.}
        \label{Fig:induction}
\end{figure}

@@

When the plants had reached maturity, we harvested the entire above-ground portion of each plant. For the outdoor sites, harvest occurred between the 1st and 5th of July 2022, while we harvested the glasshouse plants between the 19th and 21st of September 2022. We placed harvested material in labelled paper bags, and dried it a 60ºC drying oven for 48 hours. To harvest leaf material for analysis, we selected stems with flower heads, and removed the three leaves closest to the flowers. Since portions of the plant body have different silicon contents (\cite{dai_genetic_2005}), we chose a consistent set of leaves to minimize introduced variation. We picked leaves until approximately 200mg of dry leaf was collected. Some plants did not yield enough leaf tissue to meet to 200mg threshold. We pooled leaf material from within sites. For genotypes represented by three or more replicates within a site, we took a 100mg subsample of the harvested leaf material, and combined subsamples into a new sample. Overall, we were left with 359 samples across three sites, making approximately 115 useable genotypes per site. We packed dried leaf samples in 2 mL microcentrifuge tubes with three 3.2mm chrome steel grinding pellets, and ground in a tissuelyser ball mill for 60 seconds at 30 Hz. We stored the resulting leaf powder sealed until XRF analysis.

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.8\textwidth]{images/gwas_plots/svgtopng/si_manhattan_plots.png}
        \caption{Manhattan plots of the results of our GWAS analysis for leaf silicon content in \textit{Aegilops tauschii}. Red dots indicate a negative correlation, while blue dots denote a positive correlation. The top plot show the results for glasshouse environment, while the bottom two plots are the outdoor environments. The dotted line represents the bonferroni-corrected significance threshold.}
        \label{Fig:si_peak_plot}
\end{figure}

\begin{figure}[h]
        \centering
        \includegraphics[width = 0.8\textwidth]{images/gwas_plots/svgtopng/mn_manhattan_plot.png}
        \label{Fig:mn_peak_plot}
        \caption{Manhattan plots of the results of our GWAS analysis for leaf manganese content in \textit{Aegilops tauschii}. Red dots indicate a negative correlation, while blue dots denote a positive correlation. The top plot show the results for glasshouse environment, while the bottom two plots are the outdoor environments. The dotted line represents the bonferroni-corrected significance threshold.}
\end{figure}

>>>>

\section{Results}
Among the various cultivars, average uninduced silicon content ranged from 0.26\% to 0.91\% (Figure \ref{Fig:baseline_si}). We found strong support for a species effect on silicon content ($p=0.012, F = 6.08, df=3,10.3$). Oats had the lowest average silicon content at $0.34 \pm 0.02\% (\mu \pm SE)$, while wheat had the highest average silicon content at $0.76 \pm 0.05\% (\mu \pm SE)$ (Figure \ref{Fig:baseline_si}). Counter to our predictions, we failed to find strong support for inducible increases in silicon content among the tested plant species. Despite a small p-value for an induction treatment effect ($p = 0.011, F = 4.62, df =2, 186$), the only significant pairwise differences between treatments were in wheat, where silicon content was lowest in the insect treatment group (Figure \ref{Fig:induction}). We found a significant interaction between Species and Induction treatment ($p = 0.040, F = 2.52, df = 6,186$), likely driven by the stark differences between wheat and the other species' responses to the induction treatments (Figure \ref{Fig:induction}).
We found a significant effect of species, but not induction treatment, on phenolic content (Figure \ref{Fig:induction}, Table \ref{Tab:phe_params}). Oats had the highest average leaf phenolic content [$31.9 \pm 3.66 mg/kg$], while barley had the lowest average phenolic content [$12.9 \pm 1.82 mg/g$]. 
We did not find a significant relationship between silicon and phenolic content ($p = 0.15$), despite an apparent negative trend in the scatter plot (Figure \ref{Fig:phe_si_scatter}).

@@

We did not see a clear relationship between silicon and phenolic content in our plants, matching the observations of \textcite{waterman_short-term_2021}). Though they observed a negative relationship between silicon and phenolics in plants grown without supplemental silicon, this relationship disappeared when they added silicon to the growth media. As previously discussed, our study seems to have imparted moderated silicon limitation on the plants, and this may explain the hint of a negative trend observable in the scatterplot (Figure \ref{Fig:phe_si_scatter}). Our observations on silicon content variation among species complement those of \textcite{simpson_still_2017}. In their study, they tested 8 species groups of wild and domesticated cereal crops, including wheat and barley. Our results for wheat and barley silicon and phenolic content more or less follow the patterns they observed, with wheat having significantly higher silicon content, and barley having much less, while the two species have more similar phenolic content. \textcite{simpson_still_2017} did not test triticale nor oats, so this study does add first records of the silicon content and phenolic content for these species. Both of these traits are dependent on local environmental conditions (\cite{quigley_soil_2020,schaller_silicon_2012}), yet we can interpret relative differences for plants grown under common conditions, though we likely will find quantitative variation in silicon and phenolic content among studies and growing environments. In our study, oats had the highest phenolic concentration, and the lowest silicon content. The relative difference in the phenolic and silicon content of oats and wheat in our study was larger than largest difference in content in \textcite{simpson_still_2017}, but interestingly wheat is much closer to oats phylogenetically than to sorghum, the highest phenolic content species group Simpson's study. Though phylogenetic relationships are one driver of silicon and phenolic contents, there are likely proximal evolutionary (eg. domestication) and ecological (eg. growing environment) drivers that also determine the final defensive phenotype of cereal species. Understanding how these various factors constrain the range of defensive expression for these species will help bring focus to efforts to translate silicon ecology into the field of agriculture.

@@

\begin{figure}[h]
        \centering
        \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/spp_si_content.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/spp_phenolic_content.png}
        \end{subfigure}

        \centering
        \caption{Baseline (uninduced) silicon content (\% dry matter) and phenolic content (mg/kg) in the cereal cultivars used in this study. Cultivar species is indicated by point color. Error bars are 1 Standard Error.}
        \label{Fig:baseline_si}
\end{figure}

\begin{figure}[h]
        \centering
        \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/induction_plot_letters.png}
        \end{subfigure}
        \begin{subfigure}[b]{0.49\textwidth}
                \centering
                \includegraphics[width = \textwidth]{images/phenolic_induction_plot_letters.png}
        \end{subfigure}
        \caption{The response of leaf silicon content (left) and leaf phenolic content (right) for four cereal species across three defence induction treatments. We treated plants either with a 1mM methyl jasmonate spray, or exposure to house crickets (\textit{Acheta domesticus}). Leaves were sampled 18 hours after treatment, and were analysed for silicon using XRF or for phenolic content using the Fast Blue BB method. Dots and error bars are the mean $\pm$ 1 Standard Error.}
        \label{Fig:induction}
\end{figure}

@@

@@

\begin{figure}[h]
        \centering
        \includegraphics[width=0.8\textwidth]{images/gwas_plots/svgtopng/si_manhattan_plots.png}
        \caption{Manhattan plots of the results of our GWAS analysis for leaf silicon content in \textit{Aegilops tauschii}. Red dots indicate a negative correlation, while blue dots denote a positive correlation. The top plot shows the results for the glasshouse environment, while the bottom two plots are the outdoor environments. The dotted line represents the bonferroni-corrected significance threshold.}
        \label{Fig:si_peak_plot}
\end{figure}

\begin{figure}[h]
        \centering
        \includegraphics[width = 0.8\textwidth]{images/gwas_plots/svgtopng/mn_manhattan_plot.png}
        \label{Fig:mn_peak_plot}
        \caption{Manhattan plots of the results of our GWAS analysis for leaf manganese content in \textit{Aegilops tauschii}. Red dots indicate a negative correlation, while blue dots denote a positive correlation. The top plot shows the results for the glasshouse environment, while the bottom two plots are the outdoor environments. The dotted line represents the bonferroni-corrected significance threshold.}
\end{figure}
